---
title: "Multimorbidity and Access to Social Care"
subtitle: "Exploiting emerging adminsitrative datasets in Scotland"
author: "David Henderson"
date: "13th September 2018"
output:
  beamer_presentation:
    includes:
      in_header: my_styles.tex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


## Project funding and supervision

\vspace{1cm}

\begin{wideitemize}
\item Scottish Government
\item ESRC
\item UBDC
\item Nick Bailey, Colin McCowan, Stewart Mercer
\end{wideitemize}

\hspace*{8cm}\includegraphics[width = 2cm]{images/ubdc_logo.png}

## Project outline

\begin{figure}
\centerline{\includegraphics[width=0.8\textwidth, height=0.6\textheight]{images/Prev_multi_morb.png}}
\end{figure}

Barnett et al (2012)

##Project outline


\includegraphics[width=0.8\textwidth, height=0.8\textheight]{images/icl_pc.png}

##Project outline

\includegraphics[width=0.8\textwidth, height=0.8\textheight]{images/anycare_map.png}

##Project outline


In people over the age of 65 in Scotland:  

\begin{enumerate}
\item
\begin{enumerate}[(a)]
\item What are the socioeconomic, demographic, and geographic patterns in the use of social care? 
\item Is there an association between multimorbidity status and the amount and type of social care use over time? Does this vary by the patterns described in 1(a)?  
\end{enumerate}
\item 
\begin{enumerate}[(a)]
\item Is there an association in the use of social care services, multimorbidity status and unscheduled health care use?
\item Do multimorbidity status and social care use predict mortality?
\end{enumerate}
\end{enumerate}

##Project outline

###Study period 2011-2016

\includegraphics[width=0.8\textwidth]{images/linkage.png}


##Results


```{r cohort, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
`%nin%` <- negate(`%in%`)

read_csv("tab_1.csv") %>% 
  select(measure, all_years) %>% 
  mutate(all_years = sub("\\(.*","", .$all_years),
         all_years = as.integer(all_years)) %>% 
  mutate(pct = round(all_years/1104226 * 100,1)) -> cohort

x_labels <- c("Total (n)", "Female", "Male")

cohort %>% 
  filter(measure %in% c("n", "n_male", "n_female")) %>%
  ggplot(aes(reorder(measure, all_years), all_years)) +
  geom_col(fill = "#4477AA") +
  geom_text(aes(label=paste0(pct,"%")), size=4, hjust = -0.3, vjust = -0.2) +
  scale_x_discrete(labels = rev(x_labels)) +
  scale_y_continuous(limits = c(0, 1200000), 
                     breaks = scales::pretty_breaks()) +
  coord_flip() +
  theme_minimal(base_size = 16) +
  labs(title = "Count of individuals in cohort (1)",
       x = "",
       y = "") 

```


##Results

```{r plot_2}


x_labels <- c("Total (n)", "Aged 65-75", "Aged 76-85", "Aged 86+")

cohort %>% 
  filter(measure %in% c("n", "n_6575", 
                         "n_7685", "n_86plus")) %>%
  ggplot(aes(reorder(measure, all_years), all_years)) +
  geom_col(fill = "#4477AA") +
  geom_text(aes(label=paste0(pct,"%")), size=4, hjust = -0.3, vjust = -0.2) +
  scale_x_discrete(labels = rev(x_labels)) +
  scale_y_continuous(limits = c(0, 1200000), 
                     breaks = scales::pretty_breaks()) +
  coord_flip() +
  theme_minimal(base_size = 16) +
  labs(title = "Count of individuals in cohort (2)",
       x = "",
       y = "") 
```


##Results

```{r plot_3}
x_labels = c("Total (n)", "At least one\nprescription", "At lease one\nUSC episode",
           "At least one\nA & E attendance", "At least one\nunplanned admission",
           "Appeared in SCS", "Died")

cohort %>% 
  filter(measure %nin% c("n_male", "n_female", "n_6575", 
                         "n_7685", "n_86plus")) %>%
  ggplot(aes(reorder(measure, all_years), all_years)) +
  geom_col(fill = "#4477AA") +
  geom_text(aes(label=paste0(pct,"%")), size=4, hjust = -0.3, vjust = -0.2) +
  scale_x_discrete(labels = rev(x_labels)) +
  scale_y_continuous(limits = c(0, 1200000), 
                     breaks = scales::pretty_breaks()) +
  coord_flip() +
  theme_minimal(base_size = 16) +
  labs(title = "Count of individuals in cohort (3)",
       x = "",
       y = "") 
```

##Results

\includegraphics[width=0.95\textwidth]{images/plot_agesimd-1.png}


##Results

\includegraphics[width=0.95\textwidth]{images/plot_simd-1.png}

##Results

\includegraphics[width=0.95\textwidth]{images/episode_count-1.png}


##Summary

\begin{wideitemize}
\item Linkage of health and social care datasets is possible in Scotland
\item Lots still to be done in terms of quality
\item Preliminary results suggest there is \emph{not} an inverse social care law
\item Impact of Multimorbidity on social care and both of these on unscheduled care use still to be analysed
\item Watch this space....
\end{wideitemize}

##Acknowledgements

\begin{wideitemize}
\item Scottish Government and NHS National Services Scotland
\item UBDC and eDRIS for secure data management
\item Amy Tilbrook
\end{wideitemize}

##Thank you

\begin{wideitemize}
\item www.davidhen.com 
\item www.github.com/davidhen \faGithub
\item www.ubdc.ac.uk
\item @\_davidhen \faTwitter
\end{wideitemize}

\hspace*{8cm}\includegraphics[width = 2cm]{images/ubdc_logo.png}

