---
title: "Multimorbidity and Access to Social Care"
subtitle: "Exploiting emerging adminsitrative datasets in Scotland"
author: "David Henderson"
date: "3rd April 2018"
output:
  beamer_presentation:
    includes:
      in_header: my_styles.tex
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction

\begin{wideitemize}
\item  Personal Background
\item  Project funding and supervision
\item  Project outline
\item  Progress
\end{wideitemize}


## Personal Background

\begin{figure}
\centerline{\includegraphics[width=0.85\textwidth, height=0.85\textheight]{images/stressful.jpg}}
\end{figure}


##

\begin{figure}
\centerline{\includegraphics[width=0.80\textwidth, height=0.80\textheight]{images/phd.png}}
\end{figure}

<http://www.thedailymash.co.uk/news/science-technology/people-doing-phds-admit-its-just-an-excuse-to-fanny-about-20170926136446>

## Project funding and supervision

\begin{wideitemize}
\item<1-> Scottish Government
\item<2-> ESRC
\item<3-> UBDC
\item<4-> Nick Bailey, Colin McCowan, Stewart Mercer
\end{wideitemize}

## Project outline

###Background (a) Multimorbidity

\begin{figure}
\centerline{\includegraphics[width=0.8\textwidth, height=0.6\textheight]{images/Prev_multi_morb.png}}
\end{figure}

Barnett et al (2012)

##Project outline
###Background (b) What we know..

\begin{itemize}
\item<1-> There is a strong socioeconomic gradient observed for those with multimorbidity which feeds the inverse care law in primary care services.
\item<2-> MM associated with higher mortality, psychological distress, worse QOL, worse functional status, and increased health care use.
\item<3-> But what about social care?????
\end{itemize}

##Project outline
###Background (c) Social care and multimorbidity

\begin{figure}
\centerline{\includegraphics[width=0.8\textwidth, height=0.6\textheight]{images/NICE.png}}
\end{figure}

##Project outline
###Background (d) Social Care

\begin{figure}
\centerline{\includegraphics[width=0.8\textwidth, height=0.6\textheight]{images/3MT_plot.png}}
\end{figure}

\faGithub\href{https://github.com/davidhen/social_care_open_data/blob/master/rmds/3MT_plot.Rmd}{ link to code}

##Project outline
### Aims

\begin{itemize}
\item<1-> Describe and and compare social inequalities in the use of social care services using linked health and social care data. 
\item<2-> Explore the effects of social care use for those with multimorbidity on;
\begin{itemize}
\item unscheduled health care use
\item mortality
\end{itemize}
\item<3-> Assess the validity of existing administrative data on social care for research purposes. 
\end{itemize}

##Project outline
###Research questions:

In people over the age of 65 in Scotland:  

\begin{enumerate}
\item
\begin{enumerate}[(a)]
\item What are the socioeconomic, demographic, and geographic patterns in the use of social care? 
\item Is there an association between multimorbidity status and the amount and type of social care use over time? Does this vary by the patterns described in 1(a)?  
\end{enumerate}
\item 
\begin{enumerate}[(a)]
\item Is there an association in the use of social care services, multimorbidity status and unscheduled health care use?
\item Do multimorbidity status and social care use predict mortality?
\end{enumerate}
\end{enumerate}

##Project outline
###Data sources
\begin{itemize}
\item<1-> Demographics, Deaths, and SIMD (CHI database and NSS)
\item<1-> Social care survey (Scottish Government)
\item<1-> Prescribing information (ISD)
\item<1-> Unscheduled Care Data Mart (ISD)
\item<1-> SMR01, SMR04, A \& E, and USC LTC diagnoses
\end{itemize}
###Study period
\begin{itemize}
\item<2-> 1st April 2011 to 31st March 2016
\end{itemize}
###Cohort
\begin{itemize}
\item<3-> Everyone born before 31st March 1951 (over 65s)
\end{itemize}

##Project outline

```{r tribble}
library(tibble)

table <- tribble(~"", ~"n", ~"Total number of records",
                 "Total Cohort", "1,134,445", "",
                 "Total Deaths", "274,011", "",
                 "At least one prescription", "1,109,168", "134 million",
                 "Captured by social care survey", "227,345", "663,809",
                 "At least one episode USC", "845,893", "3,772,402")
knitr::kable(table)
```


##Project outline
###Measures
\begin{enumerate}
\item<1-> Multimorbidity
\begin{enumerate}[(a)]
\item Count of regular medicines
\item Count of prescribing from BNF chapters
\item Cross reference with diagnoses data
\end{enumerate}
\item<2-> Social Care
\begin{enumerate}[(a)]
\item Any social care (yes/no)
\item Weekly hours of home care
\item Telecare (yes/no)
\end{enumerate}
\item<3->Unscheduled health care use
\begin{enumerate}
\item Any USC (yes/no)
\item Count of USC episodes
\item Count of USC category (A\&E, Admissions)
\end{enumerate}
\end{enumerate}


##Progress
###Weekly variation in number of individuals receiving home care in Renfrewshire council area

\begin{figure}
\centerline{\includegraphics[width=0.8\textwidth, height=0.7\textheight]{images/renfrew.png}}
\end{figure}

##Progress

```{r timeline, message=FALSE, warning=FALSE, fig.height=4, fig.width=10}
library(ggplot2)
library(dplyr)
library(ggalt)
library(cowplot)
library(tibble)
library(lubridate)


#Create data to plot
data <- tribble( ~start_date, ~event, ~displ,
                ymd(20151001), "PhD Starts", 0.4,
                ymd(20160201), "Initial meeting with Renfrewshire", 1,
                ymd(20160524), "Ethical Approval", 0.5,
                ymd(20160712), "RAC approval", 0.3,
                ymd(20170208), "Ethical approval",-0.9,
                ymd(20170210), "RAC approval", -0.7,
                ymd(20170906), "DSA signed", 0.5,
                ymd(20170921), "Data transferred", 0.3,
                ymd(20180227), "Analysis complete", 0.3,
                ymd(20161118), "RAC submission", -1.1,
                ymd(20170404), "PBPP submission", -0.5,
                ymd(20170505), "PBPP approval", -0.3,
                ymd(20160429), "RAC submission", 0.7,
                ymd(20171122), "Health data available", -0.9,
                ymd(20171213), "DSA signed", -0.7,
                ymd(20180221), "SCS available", -0.3)


#Function to shift x-axis to 0 taken from https://stackoverflow.com/questions/39071002/moving-x-or-y-axis-together-with-tick-labels-to-the-middle-of-a-single-ggplot-n

shift_axis <- function(p, xmin, xmax, y=0){
  g <- ggplotGrob(p)
  dummy <- data.frame(y=y)
  ax <- g[["grobs"]][g$layout$name == "axis-b"][[1]]
  p + annotation_custom(grid::grobTree(ax, vp = grid::viewport(y=1, height=sum(ax$height))), 
                        ymax=y, ymin=y) +
    annotate("segment", y = 0, yend = 0, x = xmin, xend = xmax, 
             arrow = arrow(length = unit(0.1, "inches"))) +
    theme(axis.text.x = element_blank(), 
          axis.ticks.x=element_blank())

}

#Conditionally set whether text will be above or below the point
vjust = ifelse(data$displ > 0, -1, 1.5)

cond_colour <- 
  case_when(
    data$start_date ==  c("2015-10-01") ~ "black",
    data$start_date == c("2017-02-08", "2017-02-10") | data$start_date == "2016-11-18" |
      data$start_date == "2017-04-04" | data$start_date == "2017-05-05" |
      data$start_date == "2017-11-22" | data$start_date == "2017-12-13" |
      data$start_date == "2018-02-21" ~ "blue", 
    TRUE ~ "red")




#plot
p1 <- data %>% 
  ggplot(aes(start_date, displ)) +
  geom_lollipop(point.size = 2, point.colour = cond_colour) +
  geom_text(aes(x = start_date, y = displ, label = event), data = data,
            hjust = 0, vjust = vjust, size = 3) +
  theme(axis.title = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.line = element_blank(),
        axis.text.x = element_text(size = 7)) +
  expand_limits(x = c(ymd(20151201), ymd(20181001)), y = c(1.2, -0.6)) +
  scale_x_date(breaks = scales::pretty_breaks(n = 12))

#and run the function from above
timeline <- shift_axis(p1, ymd(20151001), ymd(20181001))
timeline <- timeline + 
  annotate("text", x = ymd(20160201), y = -0.6, label = "Pilot Project", colour = "red") +
  annotate("text", x = ymd(20160201), y = -0.7,   label = "Main Project", colour = "blue")
timeline + labs(caption = "RAC = Research Approvals Committee\nPBPP = Public Benefit and Privacy Panel\nDSA = Data Sharing Agreement") +
  theme(plot.caption = element_text(size = 8))

rm(list = c("data", "p1", "timeline", "vjust", "shift_axis", "table", "cond_colour"))
```

##Any Questions ??

\begin{figure}
\centerline{\includegraphics[width=0.8\textwidth, height=0.7\textheight]{images/questions.png}}
\end{figure}


