---
title: "Multimorbidity and Access to Social Care"
author: "David Henderson"
date: "26 September 2017"
output:
  beamer_presentation:
    includes:
      in_header: my_styles.tex
    highlight: monochrome
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction

\begin{wideitemize}
\item<1->  item 1
\item<2-> item 2
\item<3->  item 3
\end{wideitemize}


## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Output

```{r mtcars, echo=TRUE}
dplyr::glimpse(mtcars)
```

## Slide with table

```{r cars, results='asis'}
print(xtable::xtable(head(iris), type = "latex", booktabs = TRUE, caption = "Sample Table"),
      comment = FALSE)
```

## Slide with Plot

```{r pressure}
plot(pressure)
```